### 2021-07-07 doc/thesis/chapters/dsp.tex
nyquist done

\begin{equation}
  m_{BB}(t) = m(t) \cdot c(t)
\end{equation}

\includepicture [1.0] [0] {Signal spectrum after the down-conversion.} {down_conversion_spectrum} {img/down-conversion-AM-dsb-sc}

\section{Nyquist criteria}

in digital: any DSP must meet nyquist criteria \\
show aliasing in spectrum\\

>>>>

\begin{equation}
  m_{BB}(t) = A_m m(t) \cdot A_c c(t)
\end{equation}

Assuming that $m(t) = cos(\omega_mt)$, $c(t) = cos(\omega_ct)$ and $A_m = A_c = 1$,

\begin{equation}
  m_{BB}(t) = cos(\omega_mt) \cdot cos(\omega_ct)
\end{equation}

follows.
By applying the mathematical theorem

\begin{equation}
  cos(\alpha) \cdot cos(\beta) = \frac{1}{2} \cdot \bigl[ cos(\alpha-\beta) + cos(\alpha+\beta) \bigr]
\end{equation}

the final formula of the conversion can be represented as

\begin{equation}
 m_{BB}(t) = \frac{1}{2} \cdot \Bigl[ cos\bigl((w_m-w_c)t\bigr) + cos\bigl((w_m+w_c)t\bigr) \Bigr]
\end{equation}

The graphical representation in the frequency spectrum is shown in Fig.\ref{fig:down_conversion_spectrum_steps}.
As mentioned above, the assumption is a direct conversion to baseband, so $w_m = w_c$.
Looking at the positive original part, it is shown that only half of its energy is translated into the desired baseband.
However, half of the energy of the original negative part is also shifted to baseband.
The superposition of the two parts results in an amount of energy that is again equal to the original signal.

\includepicture [1.0] [0] {Signal spectrum, explaining the process of down-conversion.} {down_conversion_spectrum_steps} {img/down-conversion-AM-dsb-sc}

\section{Sampling theorem}

Signal processing is often done in the digital domain, using digital signals.
A digital signal is obtained by sampling an analog signal in a periodic interval, the so-called sampling frequency $f_s$.

The sampling frequency needs to meet the sampling theorem, in order to correctly represent the analog signal.
The sampling theorem, by Nyquist and Shannon mainly consists of the following formula.

\begin{equation}
  f_{s} > 2*f_{signal}
\end{equation}

In words: The sampling frequency must be higher than twice the frequency of the signals' highest frequency content.
Any digital signal processing technique must hold the specification of the sampling theorem \cite{AlessioSilviaMaria2016DSPa}.

In case the theorem is violated, the effect of aliasing occurs on the generated signal.
The sampling process introduces spectral replicas at integer multiples of the sampling frequency $f_s$.
These replicas overlap, if $f_s$ does not meet the theorem, which results in signal distortion.
Fig.\ref{fig:aliasing_freq_spectrum} illustrates the effect of aliasing in the frequency spectrum \cite{ThyagarajanK.S2019ItDS}.

\includepicture [0.9] [0] {
  Signal spectrum, explaining the process of down-conversion.\\
  Diagram A) $f_s$ is high enough and thus meets the theorem.\\
  Diagram B) $f_s$ exactly meets the theorem.\\
  Diagram C) $f_s$ is too low and thus creates overlaps of the spectral replicas - aliasing \cite{ThyagarajanK.S2019ItDS}.
} {aliasing_freq_spectrum} {img/sampling-aliasing}
